---
title: "Reproducible Research Project 1"
author: "Jeff Lee"
---

## Load required libraries

```{r}
library(knitr)
library(ggplot2)
library(dplyr)
```

## Reading activity data set

```{r}
activity <- read.csv("activity.csv", header = TRUE)
```

## Question 1 - What is the total number of steps taken per day

```{r}
stepsperday <- tapply(activity$steps, activity$date, sum)
hist(stepsperday, breaks = 50, col = "blue", xlab = "Steps", main = "Histogram Of Steps Per Day")
mean(stepsperday, na.rm = TRUE)
median(stepsperday, na.rm = TRUE)
```

## Question 2 - What is the average daily activity pattern?

```{r}
stepsbyinterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(as.numeric(names(stepsbyinterval)),stepsbyinterval, xlab = "Interval", ylab = "Steps", main = "Average Activity By Interval", type = "l")
sort(stepsbyinterval, decreasing = TRUE)[1]
```

## Question 3 - Imputing missing values

```{r}
sum(is.na(activity$steps))
```

Strategy to impute is to insert the mean of data in for the na values.

```{r}
activity2 <- activity
nadata <- is.na(activity2$steps)
avg_interval <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify = TRUE)
activity2$steps[nadata] <- avg_interval[as.character(activity2$interval[nadata])]

sum(is.na(activity2$steps))

stepsperdaynona <- tapply(activity2$steps, activity2$date, sum)
hist(stepsperdaynona, breaks = 50, col = "blue", xlab = "Steps", main = "Histogram Of Steps Per Day (Imputed Data)")
mean(stepsperdaynona, na.rm = TRUE)
median(stepsperdaynona, na.rm = TRUE)
```

Mean stays the same while median increases slightly.

## Question 4 - Are there differences in activity patterns between weekdays and weekends?

```{r}
activity2$day <- ifelse(weekdays(as.Date(activity2$date)) == "Saturday" | weekdays(as.Date(activity2$date)) == "Sunday", "Weekend", "Weekday")

stepsbyintervalweekend <- tapply(activity2[activity2$day == "Weekend" ,]$steps, activity2[activity2$day == "Weekend" ,]$interval, mean, na.rm = TRUE)

stepsbyintervalweekday <- tapply(activity2[activity2$day == "Weekday" ,]$steps, activity2[activity2$day == "Weekday" ,]$interval, mean, na.rm = TRUE)


par(mfrow=c(1,2))

plot(as.numeric(names(stepsbyintervalweekday)), 
     stepsbyintervalweekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Average Activity By Interval (Weekdays)", 
     type = "l")

plot(as.numeric(names(stepsbyintervalweekend)), 
     stepsbyintervalweekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Average Activity By Interval (Weekend)", 
     type = "l")
```
